Debugging of RxJS-based Applications

### 2020-09-29 paper.tex
Editing

Almost all participants from our interviews and war story reports showed a tendency to manually modify source code with trace logs during the hypothesis instrumentation phase when debugging RxJS code. This practice is often not perceived as efficient since the evaluation and interpretation of trace information tends to be cumbersome and very time-consuming. Also, removing log statements after a successful debugging process might leave new bugs in production code if not done carefully. We identified this technique as one of the primary debugging practices when software engineers work with RxJS-based code, which supports earlier results by Banken et al. \cite{10.1145/3180155.3180156} further.

@@

We were able to observe how all subjects predominantly used manual source code augmentation by adding trace logs, which confirmed previous results by Banken et al. \cite{10.1145/3180155.3180156}. Two of the subjects used traditional debugging utilities in order to inspect the program's state at runtime in addition. All subjects used the new information gained to refine their hypothesis about underlying problems before they started a new iteration in the debugging process. We could not observe the extraction to and reintegration from an external tool. All subjects exhibited the debugging behavior described in our hypothesis.

@@

The best RP debugging tools are useless if either the hurdle to use them is too high, or engineers do not understand which particular part of the debugging process they can benefit from them. Salvaneschi et al. \cite{10.1145/2884781.2884815} provided in their previous study on \emph{Reactive Inspector} for \emph{REScala} evidence on the effectiveness of a fully integrated RP debugging solution, which supports developers in their daily work using the Eclipse IDE. Hence, we can postulate an answer to our first research question RQ1: The most significant challenge software engineers face when debugging RxJS-based programs is to know \emph{when} they should apply \emph{what} tool to resolve their current problem in the \emph{most efficient} way.

@@

We see the biggest shortcoming of current RxJS-oriented debugging solutions like rxjs-spy, RxFiddle, or RxViz in fact that they are not integrated in established development environments (e.g., IDEs or internet browser developer tools). This leads to the practice of manually augmenting code itself rather then working with it in a less obtrusive, fully integrated way as we were able to proof in our observational study. Using specialized utilities is an extra effort an engineer has to invest every time they want to debug a data-flow: Either tagging an observable for rxjs-spy or extracting parts of it to an external environment, all of these practices require engineers to ``go the extra mile'' in order to inspect the runtime behavior of an RxJS-based application. The additional effort might be neglectable when treating a rather complex data-flow composition, but it holds back engineers from applying the tools to simple observables like in the first block of the experiment we conducted.

>>>>

Almost all participants from our interviews and war story reports showed a tendency to manually modify source code with trace logs during the hypothesis instrumentation phase when debugging RxJS code. This practice is often not perceived as efficient since the evaluation and interpretation of trace information tends to be cumbersome and very time-consuming. Also, removing log statements after a successful debugging process might leave new bugs in production code if not done carefully. Like Banken et al. \cite{10.1145/3180155.3180156} before, we identified this technique as one of the primary debugging practices when software engineers work with RxJS-based code.

@@

We were able to observe how all subjects predominantly used manual source code augmentation by adding trace logs. Two of the subjects used traditional debugging utilities in order to inspect the program's state at runtime in addition. All subjects used the new information gained to refine their hypothesis about underlying problems before starting a new iteration in the debugging process. We could not observe the extraction to and reintegration from an external tool. All subjects exhibited the debugging behavior described in our hypothesis. Further, we could verify previous results by Banken et al. \cite{10.1145/3180155.3180156} successfully as well.

@@

The best RP debugging tools are useless if either the hurdle to use them is too high, or engineers do not understand which particular part of the debugging process they can benefit from them. Salvaneschi et al. \cite{10.1145/2884781.2884815} provided in their previous study on the \emph{Reactive Inspector} for \emph{REScala} evidence on the effectiveness of a fully integrated RP debugging solution, which supports developers in their daily work using the Eclipse IDE. Hence, we can postulate an answer to our first research question RQ1: The most significant challenge software engineers face when debugging RxJS-based programs is to know \emph{when} they should apply \emph{what} tool to resolve their current problem in the \emph{most efficient} way.

@@

We see the biggest shortcoming of current RxJS-oriented debugging solutions like rxjs-spy, RxFiddle, or RxViz in fact that they are not integrated in established development environments (e.g., IDEs or internet browser developer tools). This leads to the practice of manually augmenting code itself rather then working with it in a less obtrusive, fully integrated way as we were able to proof in our observational study. Using specialized utilities is an extra effort an engineer has to invest every time they want to debug a data-flow: Either tagging an observable for rxjs-spy or extracting parts of it to an external environment, all of these practices require engineers to ``go the extra mile'' in order to inspect the runtime behavior of an RxJS-based application. The additional effort might be neglectable when treating a rather complex data-flow composition. However, it holds back engineers from applying the tools to simple observables like in the first block of the experiment we conducted.
